---
title: "Reproducible Research: Peer Assessment 1"
author: "Daniel Hicks"
date: "Saturday, June 13, 2015"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
#Load libraries:
library(ggplot2)
library(dplyr)
library(lubridate)
#Load data:
activity<-read.csv('activity.csv')
```
## What is mean total number of steps taken per day?
Calculate the total steps:

```{r}
totals<-summarise_each(group_by(na.omit(activity), date), funs(sum))
totals
```

Create a histogram of the total number of steps taken each day:
```{r}
hist(totals$steps, main = 'Total Steps', col = 'orange')
```

Calculate and report the mean of the total number of steps taken per day:
```{r}
mean(totals$steps)
```
Calculate and report the median of the total number of steps taken per day:
```{r}
median(totals$steps)
```
## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

```{r}
stepinterval<-summarise_each(group_by(na.omit(activity), interval), funs(mean))
plot(x=stepinterval$interval,y=stepinterval$steps, type = "l", col='orange', ylab='average number of steps taken')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?:
```{r}
stepinterval[which.max(stepinterval$steps), ]$interval
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(activity$steps))
```
Replace each NA entry with the mean for the corresponding 5-minute interval.Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
fullact<- activity %>% group_by(interval) %>% mutate(steps= replace(steps, is.na(steps), mean(steps, na.rm=TRUE)))

```
Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

Calculate the total steps:

```{r}
totalsnew<-summarise_each(group_by(fullact, date), funs(sum))
totalsnew
```

Create a histogram of the total number of steps taken each day:
```{r}
hist(totalsnew$steps, main = 'Total Steps', col = 'lightblue')
```

Calculate and report the mean of the total number of steps taken per day:
```{r}
mean(totalsnew$steps)
```
Calculate and report the median of the total number of steps taken per day:
```{r}
median(totalsnew$steps)
```
Do these values differ from the estimates from the first part of the assignment? 
```{r}
mean(totals$steps) - mean(totalsnew$steps)

median(totals$steps) - median(totalsnew$steps)
```
The impact of replacing the NA entries with the mean for the 5-minute interval, is a slight ajustment to the median to equal the mean.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
dotw <- weekdays(as.Date(fullact$date))
dotw <- ifelse(test = dotw %in% c("Saturday", "Sunday"), yes="weekend", "weekday")
fullact$dotw <- as.factor(dotw)
head(fullact)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
ggplot(fullact)+geom_line(aes(fullact$interval,fullact$steps),colour='orange', alpha=0.8)+facet_wrap(~dotw,ncol=1)+xlab('Intervals')+ylab('Number of steps')
```
